Computational Methods for Principle Component Analysis Applied to Genotype Data

### 2018-06-20 doc/paper/paper.tex
Add all the plots to paper

\begin{figure}[p]
\begin{subfigure}{.25\textwidth}
  \centering
  \includegraphics[width=.98\linewidth]{img/ggsim1000_100000_200_2_1_100_1_sturef_ggsim1000_100000_1000_2_1_100_0_sp}
\end{subfigure}%
\begin{subfigure}{.25\textwidth}
  \centering
  \includegraphics[width=.98\linewidth]{img/ggsim1000_100000_200_2_1_100_1_sturef_ggsim1000_100000_1000_2_1_100_0_ap}
\end{subfigure}%
\begin{subfigure}{.25\textwidth}
  \centering
  \includegraphics[width=.98\linewidth]{img/ggsim1000_100000_200_2_1_100_1_sturef_ggsim1000_100000_1000_2_1_100_0_oadp}
\end{subfigure}%
\begin{subfigure}{.25\textwidth}
  \centering
  \includegraphics[width=.98\linewidth]{img/ggsim1000_100000_200_2_1_100_1_sturef_ggsim1000_100000_1000_2_1_100_0_adp}
\end{subfigure}
\hfill
\begin{subfigure}{.25\textwidth}
  \centering
  \includegraphics[width=.98\linewidth]{img/ggsim1500_100000_200_2_1_100_1_sturef_ggsim1500_100000_1500_2_1_100_0_sp}
\end{subfigure}%
\begin{subfigure}{.25\textwidth}
  \centering
  \includegraphics[width=.98\linewidth]{img/ggsim1500_100000_200_2_1_100_1_sturef_ggsim1500_100000_1500_2_1_100_0_ap}
\end{subfigure}%
\begin{subfigure}{.25\textwidth}
  \centering
  \includegraphics[width=.98\linewidth]{img/ggsim1500_100000_200_2_1_100_1_sturef_ggsim1500_100000_1500_2_1_100_0_oadp}
\end{subfigure}%
\begin{subfigure}{.25\textwidth}
  \centering
  \includegraphics[width=.98\linewidth]{img/ggsim1500_100000_200_2_1_100_1_sturef_ggsim1500_100000_1500_2_1_100_0_adp}
\end{subfigure}
\hfill
\begin{subfigure}{.25\textwidth}
  \centering
  \includegraphics[width=.98\linewidth]{img/ggsim2000_100000_200_2_1_100_1_sturef_ggsim2000_100000_2000_2_1_100_0_sp}
\end{subfigure}%
\begin{subfigure}{.25\textwidth}
  \centering
  \includegraphics[width=.98\linewidth]{img/ggsim2000_100000_200_2_1_100_1_sturef_ggsim2000_100000_2000_2_1_100_0_sp}
\end{subfigure}%
\begin{subfigure}{.25\textwidth}
  \centering
  \includegraphics[width=.98\linewidth]{img/ggsim2000_100000_200_2_1_100_1_sturef_ggsim2000_100000_2000_2_1_100_0_oadp}
\end{subfigure}%
\begin{subfigure}{.25\textwidth}
  \centering
  \includegraphics[width=.98\linewidth]{img/ggsim2000_100000_200_2_1_100_1_sturef_ggsim2000_100000_2000_2_1_100_0_adp}
\end{subfigure}
\hfill
\begin{subfigure}{.25\textwidth}
  \centering
  \includegraphics[width=.98\linewidth]{img/ggsim2500_100000_200_2_1_100_1_sturef_ggsim2500_100000_2500_2_1_100_0_sp}
\end{subfigure}%
\begin{subfigure}{.25\textwidth}
  \centering
  \includegraphics[width=.98\linewidth]{img/ggsim2500_100000_200_2_1_100_1_sturef_ggsim2500_100000_2500_2_1_100_0_ap}
\end{subfigure}%
\begin{subfigure}{.25\textwidth}
  \centering
  \includegraphics[width=.98\linewidth]{img/ggsim2500_100000_200_2_1_100_1_sturef_ggsim2500_100000_2500_2_1_100_0_oadp}
\end{subfigure}%
\begin{subfigure}{.25\textwidth}
  \centering
  \includegraphics[width=.98\linewidth]{img/ggsim2500_100000_200_2_1_100_1_sturef_ggsim2500_100000_2500_2_1_100_0_adp}
\end{subfigure}
\hfill
\begin{subfigure}{.25\textwidth}
  \centering
  \includegraphics[width=.98\linewidth]{img/ggsim3000_100000_200_2_1_100_1_sturef_ggsim3000_100000_3000_2_1_100_0_sp}
\end{subfigure}%
\begin{subfigure}{.25\textwidth}
  \centering
  \includegraphics[width=.98\linewidth]{img/ggsim3000_100000_200_2_1_100_1_sturef_ggsim3000_100000_3000_2_1_100_0_sp}
\end{subfigure}%
\begin{subfigure}{.25\textwidth}
  \centering
  \includegraphics[width=.98\linewidth]{img/ggsim3000_100000_200_2_1_100_1_sturef_ggsim3000_100000_3000_2_1_100_0_oadp}
\end{subfigure}%
\begin{subfigure}{.25\textwidth}
  \centering
  \includegraphics[width=.98\linewidth]{img/ggsim3000_100000_200_2_1_100_1_sturef_ggsim3000_100000_3000_2_1_100_0_adp}
\end{subfigure}
\caption{
  The PC scores predicted by each of the four methods
  as the reference size varies.
  Reference samples are in colors
  and study samples are in black.
  The number of variants is 100,000,
  and the study sample size is 200.
  Only the top 4 PCs are calculated.
  The data is simulated by the GGS software \citep{mathieson2012differential}.
}
\end{figure}

@@

\begin{table}[h]
  \centering
  \begin{tabular}{|l|l|l|l|l|l|}
    \hline
    Reference Size & 1000 & 1500 & 2000 & 2500 & 3000 \\
    \hline
    \multicolumn{6}{|c|}{Study Runtime} \\
    \hline
    SP         &  17  & 17   & 16   & 18   & 18   \\
    AP         &  18  & 16   & 18  & 16   & 17  \\
    OADP       &  145  & 146   & 147   & 148   & 149   \\
    ADP        &  379  & 809 & 1526 & 2556 & 4062 \\
    \hline
    \multicolumn{6}{|c|}{Error form ADP} \\
    \hline
    SP         &  188  & 139   & 112   & 91   & 84   \\
    AP         &  31  & 27   & 12  & 11   & 14  \\
    OADP       &  21  & 16   & 12   & 9   & 10   \\
    ADP        &  0  & 0 & 0 & 0 & 0 \\
    \hline
  \end{tabular}
  \caption{
    The runtimes and errors of the four methods
    as the reference size increases.
    Errors are measured as the Euclidean distance of
    The number of variants is 100,000,
    and the study sample size is 200.
    Only the top 4 PCs are calculated.
    The data is simulated by the GGS software \citep{mathieson2012differential}.
    }
\end{table}

\begin{figure}[h]

  \begin{subfigure}{.5\textwidth}
    \centering
    \includegraphics[width=.98\linewidth]{img/ggsim_runtimes.png}
  \end{subfigure}
  \begin{subfigure}{.5\textwidth}
    \centering
    \includegraphics[width=.98\linewidth]{img/ggsim_errors.png}
  \end{subfigure}
  \caption{
    The runtimes (in seconds) and errors (Euclidean distance from ADP's PC scores) for the four methods
    as the reference size increases.
    The number of variants is 100,000,
    and the study sample size is 200.
    Only the top 4 PCs are calculated.
    The data is simulated by the GGS software \citep{mathieson2012differential}.
  }
\end{figure}

\begin{table}[h]
  \centering
  \begin{tabular}{|r|r|r|r|r|}
    \hline
    Method & \multicolumn{2}{|c|}{Runtime} & \multicolumn{2}{|c|}{Error} \\
    \hline
    & Global & European & Global & European \\
    \hline
    SP         &  62   & 48  & 92  & 450 \\ 
    AP         &  65   & 48  & 48  & 33  \\
    OADP       &  363  & 349 & 10  & 32  \\
    ADP        &  6739 & 518 & 0   & 0   \\
    \hline
  \end{tabular}
  \caption{
    The study runtimes (in seconds) and erros for the four methods
    for 500 randomly selected study samples
    in the UK Biobank data.
    Errors are measured as the Euclidean distance of
    the PC scores predicted by each method
    to those predicted by ADP.
    The 2492 samples from the 1000 Genomes data
    are used as the reference.
    The number of variants in common is 145,282.
    Only the top 4 PCs are calculated.
    }
\end{table}

@@

>>>>

\begin{figure}[p]
\begin{subfigure}{.5\textwidth}
  \centering
  \includegraphics[width=.98\linewidth]{img/ggsim1000_100000_200_2_1_100_1_sturef_ggsim1000_100000_1000_2_1_100_0_sp}
\end{subfigure}%
\begin{subfigure}{.5\textwidth}
  \centering
  \includegraphics[width=.98\linewidth]{img/ggsim1000_100000_200_2_1_100_1_sturef_ggsim1000_100000_1000_2_1_100_0_ap}
\end{subfigure}
\hfill
\begin{subfigure}{.5\textwidth}
  \centering
  \includegraphics[width=.98\linewidth]{img/ggsim1000_100000_200_2_1_100_1_sturef_ggsim1000_100000_1000_2_1_100_0_oadp}
\end{subfigure}%
\begin{subfigure}{.5\textwidth}
  \centering
  \includegraphics[width=.98\linewidth]{img/ggsim1000_100000_200_2_1_100_1_sturef_ggsim1000_100000_1000_2_1_100_0_adp}
\end{subfigure}
\end{figure}

\begin{figure}[p]
\begin{subfigure}{.5\textwidth}
  \centering
  \includegraphics[width=.98\linewidth]{img/ggsim1500_100000_200_2_1_100_1_sturef_ggsim1500_100000_1500_2_1_100_0_sp}
\end{subfigure}%
\begin{subfigure}{.5\textwidth}
  \centering
  \includegraphics[width=.98\linewidth]{img/ggsim1500_100000_200_2_1_100_1_sturef_ggsim1500_100000_1500_2_1_100_0_ap}
\end{subfigure}
\hfill
\begin{subfigure}{.5\textwidth}
  \centering
  \includegraphics[width=.98\linewidth]{img/ggsim1500_100000_200_2_1_100_1_sturef_ggsim1500_100000_1500_2_1_100_0_oadp}
\end{subfigure}%
\begin{subfigure}{.5\textwidth}
  \centering
  \includegraphics[width=.98\linewidth]{img/ggsim1500_100000_200_2_1_100_1_sturef_ggsim1500_100000_1500_2_1_100_0_adp}
\end{subfigure}
\end{figure}

\begin{figure}[p]
\begin{subfigure}{.5\textwidth}
  \centering
  \includegraphics[width=.98\linewidth]{img/ggsim2000_100000_200_2_1_100_1_sturef_ggsim2000_100000_2000_2_1_100_0_sp}
\end{subfigure}%
\begin{subfigure}{.5\textwidth}
  \centering
  \includegraphics[width=.98\linewidth]{img/ggsim2000_100000_200_2_1_100_1_sturef_ggsim2000_100000_2000_2_1_100_0_ap}
\end{subfigure}
\hfill
\begin{subfigure}{.5\textwidth}
  \centering
  \includegraphics[width=.98\linewidth]{img/ggsim2000_100000_200_2_1_100_1_sturef_ggsim2000_100000_2000_2_1_100_0_oadp}
\end{subfigure}%
\begin{subfigure}{.5\textwidth}
  \centering
  \includegraphics[width=.98\linewidth]{img/ggsim2000_100000_200_2_1_100_1_sturef_ggsim2000_100000_2000_2_1_100_0_adp}
\end{subfigure}
\end{figure}

\begin{figure}
\begin{subfigure}{.5\textwidth}
  \centering
  \includegraphics[width=.98\linewidth]{img/ggsim2500_100000_200_2_1_100_1_sturef_ggsim2500_100000_2500_2_1_100_0_sp}
\end{subfigure}%
\begin{subfigure}{.5\textwidth}
  \centering
  \includegraphics[width=.98\linewidth]{img/ggsim2500_100000_200_2_1_100_1_sturef_ggsim2500_100000_2500_2_1_100_0_ap}
\end{subfigure}
\hfill
\begin{subfigure}{.5\textwidth}
  \centering
  \includegraphics[width=.98\linewidth]{img/ggsim2500_100000_200_2_1_100_1_sturef_ggsim2500_100000_2500_2_1_100_0_oadp}
\end{subfigure}%
\begin{subfigure}{.5\textwidth}
  \centering
  \includegraphics[width=.98\linewidth]{img/ggsim2500_100000_200_2_1_100_1_sturef_ggsim2500_100000_2500_2_1_100_0_adp}
\end{subfigure}
\end{figure}

\begin{figure}
\begin{subfigure}{.5\textwidth}
  \centering
  \includegraphics[width=.98\linewidth]{img/ggsim3000_100000_200_2_1_100_1_sturef_ggsim3000_100000_3000_2_1_100_0_sp}
\end{subfigure}%
\begin{subfigure}{.5\textwidth}
  \centering
  \includegraphics[width=.98\linewidth]{img/ggsim3000_100000_200_2_1_100_1_sturef_ggsim3000_100000_3000_2_1_100_0_sp}
\end{subfigure}
\hfill
\begin{subfigure}{.5\textwidth}
  \centering
  \includegraphics[width=.98\linewidth]{img/ggsim3000_100000_200_2_1_100_1_sturef_ggsim3000_100000_3000_2_1_100_0_oadp}
\end{subfigure}%
\begin{subfigure}{.5\textwidth}
  \centering
  \includegraphics[width=.98\linewidth]{img/ggsim3000_100000_200_2_1_100_1_sturef_ggsim3000_100000_3000_2_1_100_0_adp}
\end{subfigure}
\caption{
  The PC scores predicted by each of the four methods
  as the reference size varies.
  Reference samples are in colors
  and study samples are in black.
  The stars mark the centers of mass (geographic centers) of each population for the reference and study samples.
  The number of variants is 100,000,
  and the study sample size is 200.
  Only the top 4 PCs are calculated.
  The data is simulated by the GGS software \citep{mathieson2012differential}.
}
\end{figure}

@@

\begin{table}[h]
  \centering
  \begin{tabular}{|l|l|l|l|l|l|}
    \hline
    Reference Size & 1000 & 1500 & 2000 & 2500 & 3000 \\
    \hline
    \multicolumn{6}{|c|}{Study Runtime} \\
    \hline
    SP         &  17  & 17   & 16   & 18   & 18   \\
    AP         &  18  & 16   & 18  & 16   & 17  \\
    OADP       &  145  & 146   & 147   & 148   & 149   \\
    ADP        &  379  & 809 & 1526 & 2556 & 4062 \\
    \hline
    \multicolumn{6}{|c|}{Distance from ADP's results} \\
    \hline
    SP         &  188  & 139   & 112   & 91   & 84   \\
    AP         &  31  & 27   & 12  & 11   & 14  \\
    OADP       &  21  & 16   & 12   & 9   & 10   \\
    ADP        &  0  & 0 & 0 & 0 & 0 \\
    \hline
    \multicolumn{6}{|c|}{Distance from reference centers} \\
    \hline
    SP         &  335 & 161  & 113 & 71  & 60  \\
    AP         &  31  & 19   & 17  & 25  & 15  \\
    OADP       &  12  & 17   & 11  & 20  & 12  \\
    ADP        &  20  & 14   & 14  & 23  & 14  \\
    \hline
  \end{tabular}
  \caption{
    The runtimes and errors of the four methods
    as the reference size increases.
    Distances are measured by the Euclidean distance.
    Reference centers are the centers of mass of the populations
    in the reference samples.
    The number of variants is 100,000,
    and the study sample size is 200.
    Only the top 4 PCs are calculated.
    The data is simulated by the GGS software \citep{mathieson2012differential}.
    }
\end{table}

\begin{figure}[h]
  \centering
  \includegraphics[width=.98\linewidth]{img/ggsim_runtimes.png}
  \centering
  \includegraphics[width=.98\linewidth]{img/ggsim_errors_adp.png}
  \centering
  \includegraphics[width=.98\linewidth]{img/ggsim_errors_ctr.png}
  \caption{
    The runtimes and errors of the four methods
    as the reference size increases.
    Runtimes are measured in seconds.
    Errors from ADP are the Euclidean distance
    from each method's result to ADP's result.
    Errors between centers are the Euclidean distance
    from the centers of mass of the reference populations
    to the corresponding centers of mass of the study populations.
    The number of variants is 100,000,
    and the study sample size is 200.
    Only the top 4 PCs are calculated.
    The data is simulated by the GGS software \citep{mathieson2012differential}.
  }
\end{figure}

\begin{table}[H]
  \centering
  \begin{tabular}{|r|r|r|r|r|r|r|}
    \hline
    Method & \multicolumn{2}{|c|}{Runtime} & \multicolumn{2}{|c|}{Error (ADP)} & \multicolumn{2}{|c|}{Error (centers)} \\
    \hline
    & Global & European & Global & European & Global & European \\
    \hline
    SP         &  62   & 48  & 92  & 450 & 3156 & 2304 \\ 
    AP         &  65   & 48  & 48  & 33  & 2486 & 706  \\
    OADP       &  363  & 349 & 10  & 32  & 2575 & 665  \\
    ADP        &  6739 & 518 & 0   & 0   & 2578 & 649  \\
    \hline
  \end{tabular}
  \caption{
    The study runtimes (in seconds) and erros for the four methods
    for 500 randomly selected study samples
    in the UK Biobank data.
    Errors (ADP) are the Euclidean distances from
    the PC scores predicted by each method
    to those predicted by ADP.
    Errors (centers) are the Euclidean distances
    from the centers of mass of the reference samples
    to the centers of mass of the study samples.
    For the global study,
    2492 samples from the 1000 Genomes data are used as the reference.
    For the European study,
    only the 498 European samples among them are used.
    The number of variants in common is 145,282.
    Only the top 4 PCs are calculated.
    }
\end{table}

@@

\begin{figure}[p]
  \centering
  \includegraphics[width=0.98\textwidth]{img/ukb_snpscap_kgn_bial_orphans_nchunks100_sturef_kgn_bial_orphans_snps_ukb_snpscap_ukb_sp.png}
  \includegraphics[width=0.98\textwidth]{img/ukb_snpscap_kgn_bial_orphans_nchunks100_sturef_kgn_bial_orphans_snps_ukb_snpscap_ukb_ap.png}
  \includegraphics[width=0.98\textwidth]{img/ukb_snpscap_kgn_bial_orphans_nchunks100_sturef_kgn_bial_orphans_snps_ukb_snpscap_ukb_oadp.png}
  \caption{All the 488366 samples in the UK Biobank data.}
\end{figure}

\begin{figure}[p]
  \centering
  \includegraphics[width=0.98\textwidth]{img/ukb_snpscap_kgn_bial_orphans_pred_EUR_nchunks100_sturef_kgn_bial_orphans_snps_ukb_snpscap_ukb_EUR_sp.png}
  \includegraphics[width=0.98\textwidth]{img/ukb_snpscap_kgn_bial_orphans_pred_EUR_nchunks100_sturef_kgn_bial_orphans_snps_ukb_snpscap_ukb_EUR_ap.png}
  \includegraphics[width=0.98\textwidth]{img/ukb_snpscap_kgn_bial_orphans_pred_EUR_nchunks100_sturef_kgn_bial_orphans_snps_ukb_snpscap_ukb_EUR_oadp.png}
  \caption{All the 462284 European samples in the UK Biobank data.}
\end{figure}
